Fast Construction of Inter-Object Space Representations

### 2017-04-03 paper.tex
Optimality, Nate changes

\documentclass[final,3p,times,twocolumn]{elsarticle}

@@

@@

\title{Fast Construction of Inter-Object Spacing Representations}
\author{Nathan Morrical}
%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author{John Edwards}

\address{Department of Informatics and Computer Science, Idaho State University, USA}

@@

\begin{abstract}
We present a parallel quadtree algorithm that resolves between geometric objects, modeling space between objects rather than the objects themselves. Our quadtree has the property that no cell intersects more than one labeled object. A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori or because the number of cells required to resolve closely spaced objects exceeds available memory. Previous parallel algorithms that are spatially adaptive, discretizing finely only where needed, \red{Hierarchical kernel envokation isn't necessarily bad. We might move to that to remove some of the Q linear complexity overhead.} either separate points only, or make no guarantees of object separation. Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time). We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D. Our results show significant improvement over the current state of the art.

@@

%-------------------------------------------------------------------------------
% introduction
%-------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
\red{Reworded this paragraph, since the last sentence contained duplicate information provided in the first sentence.} Constructing quadtrees on objects is an important task with applications in collision detection, distance fields, generalized Voronoi Diagram (GVD) construction \cite{edwards2015approximating}, robot navigation, shape modeling, object description, and other applications. Quadtrees built on objects most often model the objects themselves, providing a space-efficient representation of arbitrarily complex objects. In other cases, quadtrees are purposed for fast retrieval as is often the case in hierarchical subdivisions of point data. However, our work centers on using quadtrees to separate, or resolve, collections of closely spaced objects, i.e., to construct a discretization such that no cell intersects more than one object.

\red{reworded the first sentence a bit.} This object separation is of some use in 2D (e.g. path planning), but becomes a very important problem in 3D. Hierarchically subdividing non-point data in a principled parallel way is surprisingly complex, and this paper lays the groundwork for our continuing work in 3D.

Modeling inter-object spacing is computationally straightforward when that spacing is large compared to the world bounding box. Approaches typically involve a uniform grid of the space, which leads to efficient computation that often uses graphics processors.

Difficulties arise when objects are close together relative to the size of the domain. An approach using a uniform grid would have excessive memory requirements in order to resolve between objects because the uniformly sized grid cell must be small enough to fit between objects at every location in the domain. Thus, an adaptive approach must be used for datasets of closely spaced objects. \red{added carriage return here}

@@

This paper extends the work done by Edwards \etal \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm that is adaptive and independent of object distribution. Our algorithm, which is targeted for the GPU, performs an order of magnitude faster than the previous work and will be an important base for later distance transform and generalized Voronoi diagram computation.

@@

@@

\paragraph{Parallel} Many recent works on fast quadtree construction using the GPU are limited to point sites \cite{bedorf2012sparse,karras2012maximizing,zhou2011data}. Kim and Liu's work~\shortcite{kim2014exact} is similar, computing the quadtree on the barycenters of triangles, giving an approximation of an object-resolving quadtree. Most quadtree approaches that support surfaces \cite{baert2013out,crassin2009gigavoxels,laine2011efficient,lefebvre2007compressed} are designed for efficient rendering, and actual construction of the quadtree is implemented on the CPU. Two works \cite{bastos2008gpu,park2010cuda} implement Adaptive Distance Fields in parallel on quadtrees but building the quadtree itself is done sequentially.  Yin \etal~\shortcite{yin2011fast} compute the octree entirely on the GPU using a bottom-up approach by initially subdividing into a complete quadtree, resulting in memory usage that is no better than using a uniform grid. We have found no GPU quadtree construction method that is fully adaptive and can resolve between objects.

@@

\subsection{Prune the octree}
During the initial quadtree construction, we can indirectly prune the initial quadtree to simplify conflict detection and reduce our memory footprint. Assume we have a numeric vertex labeling such that each vertex is labeled to match the object it belongs to. The initial binary radix tree (BRT) provided by Karras serves as a bounding volume hierarchy, and is used to generate the initial quadtree by seperating vertices regardless of their label. Since our objective is to resolve between objects of different labels, we can proactively prune the initial BRT, and subsequently the initial quadtree (see figure \ref{fig:pruning}) such that a leaf node can contain multiple vertices as long as they all have the same label.

\red{I tried to be a bit more verbose here and in the paragraph above. Let me know if anything in particular needs more clarification.} To prune the initial BRT efficiently, we label each BRT node using the following criterion: if the current node is a leaf node which seperates two vertices with identical labels, label the current node to match the label of the vertices being seperated. If the current node is a leaf node that seperates two vertices having mismatching colors, label the current node as -2, i.e. "required". Lastly, if the current node is an internal node, i.e. the current node is an ancestor of another BRT node, mark the current node as -1, i.e. "unknown''. This initial step can be done immediately after Karras' BRT construction without the need to invoke an additional kernel.

We then propagate the BRT labels up the tree in parallel, marking "unknown'' nodes as "required" when the labels of the current node's two children nodes mismatch.  Finally, we generate quadtree nodes from only the required internal binary radix tree nodes.

@@

We define a bounding cell (BCell) to be the smallest internal quadtree node which entirely contains a given facet. Given a facet defined by $n$ endpoints $P=\{p_1, p_2, \dots, p_n\}$, the quadtree address of the BCell is contained in the longest common prefix (LCP) of the Morton codes of the points in $P$. If a given LCP is more specific than any quadtree node, ie the LCP lies within a quadtree leaf, we simply take the quadtree address of the leaf that the LCP lies within. This is often the case with pruned quadtrees where entire facets may lie within a quadtree leaf. \red{This paragraph had some incorrect conclusions, so I tried to clarify a bit more.} Figure \ref{fig:data-structures-1} gives the addresses of the BCells of the facets in figure \ref{fig:conflict-find-4}.

@@

Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (figure \ref{fig:data-structures-2}).% \texttt{BCells} array construction and sorting is done in parallel with the initial Karras octree construction.

@@

    \node[anchor=south west,inner sep=0,outer sep=0] at (0,-4) {
%      \subfloat[][]{
%        \label{fig:conflict-find-2}
%        \includegraphics[width=0.44\columnwidth]{conflict-find-2.pdf} } \hfill
%      \subfloat[][]{
%        \label{fig:conflict-find-4}
%        \includegraphics[width=0.44\columnwidth]{conflict-find-4.pdf} } \hfill
      \subfloat[][]{
        \label{fig:conflict-find-5}
        \includegraphics[width=0.44\columnwidth]{conflict-find-5.pdf} }
      \subfloat[][]{
        \label{fig:conflict-find-6}
        \includegraphics[width=0.44\columnwidth]{conflict-find-6.pdf} }
    };

@@

    % two column, .4 size
    % small rectangle
    \draw[red,thick] (0.0,1.65) rectangle (2.33,4.0);
%    % upper-right line
%    \draw[red,thin] (3.27,2.25) -- (3.9,2.25);
%    % lower-left line
%    \draw[red,thin] (2.0,1.05) -- (3.9,0.4);
%    % large rectangle
%    \draw[red,thick] (3.9,0.4) rectangle (5.8,2.25);
  \end{tikzpicture}
%      \subfloat[][]{
%        \label{fig:data-structures-1}
%        \includegraphics[width=0.38\columnwidth]{data-structures-1.pdf} } \hfill
%      \subfloat[][]{
%        \label{fig:data-structures-2}
%        \includegraphics[width=0.38\columnwidth]{data-structures-3.pdf} } \hfill
%      \subfloat[][]{
%        \label{fig:data-structures-3}
%        \includegraphics[width=0.38\columnwidth]{data-structures-4.pdf} }
  \caption{
%    \protect\subref{fig:conflict-find-1} We have three objects, blue, red, and green with facets labeled A-I.
    We have three objects, blue, red, and green with facets labeled A-I.
    \protect\subref{fig:conflict-find-2} Initial vertex quadtree. \red{This vertex quadtree is unpruned. Can we prune it for consistency?}
    \protect\subref{fig:conflict-find-4} Zoomed-in to the region outlined by red in \protect\subref{fig:conflict-find-2} showing the boundary cell (BCell) computation for each facet. %These pairs are given in figure \protect\subref{fig:data-structures-1}.
    \protect\subref{fig:conflict-find-5} Conflict cells, which intersect more than one object, are highlighted.
    \protect\subref{fig:conflict-find-6} The new quadtree after conflict resolution.
%    \protect\subref{fig:data-structures-1} The bounding cells (BCells) are stored in an array initially sorted on facet index (letters are used here for clarity). The quadtree array elements are structures which store child and parent pointers (``C/P'' in the figure).
%    \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access. We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given octree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
%\protect\subref{fig:data-structures-3} For a given octree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
  }
  \label{fig:conflict-find}
\end{figure}

\begin{figure}
  \centering
      \subfloat[][]{
        \label{fig:data-structures-1}
        \includegraphics[width=0.32\columnwidth]{data-structures-1.pdf} }
      \subfloat[][]{
        \label{fig:data-structures-2}
        \includegraphics[width=0.32\columnwidth]{data-structures-3.pdf} }
      \subfloat[][]{
        \label{fig:data-structures-3}
        \includegraphics[width=0.32\columnwidth]{data-structures-4.pdf} }
  \caption{
    \protect\subref{fig:data-structures-1} The bounding cells (BCells) are stored in an array initially sorted on facet index (letters are used here for clarity). The quadtree array elements are structures which store child and parent pointers (``C/P'' in the figure).
    \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access. We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given octree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
\protect\subref{fig:data-structures-3} For a given octree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
  }
  \label{fig:data-structures}
\end{figure}

Then we use the \texttt{BCells} array and octree data structure to find the conflict cells using algorithm \ref{alg:find-conflict-cells}. We process each leaf cell $L$ in parallel (line \ref{alg:L}). First, we set $L$'s color to $-1$ (uninitialized). We then traverse each ancestor $A$ of $L$ (line \ref{alg:A}) by using the \texttt{Parent} field in the octree data structure. Using the \texttt{FFacet} and \texttt{LFacet} fields, we find, respectively, the first and (inclusive) last of possibly multiple facets bounded by $A$ (line \ref{alg:i}). The \texttt{FacetMap} array is used to find all facets bounded by bounding cell $A$ (line \ref{alg:f}). Any facet $f$ for which $A$ is the bounding cell could potentially intersect the leaf cell $L$. We test for intersection between $f$ and $L$ and store the first two facets of differing color (lines \ref{alg:7}-\ref{alg:16}). If at the conclusion of execution $L$.color is equal to $-2$ then $L$ is a conflict cell and must be resolved.

@@

\begin{figure}
  \centering
  \subfloat[][]{
    \label{fig:conflict-resolution-x}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-x.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-even}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-even.pdf} }\\
  \subfloat[][]{
    \label{fig:conflict-resolution-all}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-all.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-octree}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-octree.pdf} }\\
  \subfloat[][]{
    \label{fig:conflict-resolution-adjacent-even}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-adjacent-even.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-adjacent-all}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-adjacent-all.pdf} }
  \caption{
    \protect\subref{fig:conflict-resolution-x} A conflict cell with two lines from different objects.
    \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least one sample (red dots).
    \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least two samples. This ensures that a quadtree built from the samples using Karras' algorithm (panel \protect\subref{fig:conflict-resolution-octree}) will have no leaf cells that intersect both lines, ensuring that the new quadtree is locally free of conflict cells.
  }
  \label{fig:conflict-resolution}
\end{figure}

@@

\begin{align}
q(t) &= q = q_0 + tv \label{eqn:q} \\
r(f) &= r = r_0 + fw \label{eqn:r}
\end{align}
along with a line
\begin{align}
p(s) &= p = p_0 + su \label{eqn:p}
\end{align}
that bisects $q$ and $r$. Our strategy will be to sample points $P$ on $p(s)$ (figure \ref{fig:conflict-resolution-octree}) such that a quadtree built on $S \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent cell of $c$ will intersect both $q$ and $r$. We do this by ensuring that $P$ is sampled such that every box that intersects both $q$ and $r$ also intersects at least two points in $P$. Because Karras' algorithm guarantees that every leaf cell intersects at most one point, we know that no leaf cell will intersect $q$ and $r$ and thus no leaf cell will be a conflict cell. We will find a series of boxes such that each box's left-most intersection with $p(s)$ is a sample point meeting the above criterion. In the following discussion, $p^x$ and $p^y$ refer to the $x$ and $y$ coordinates of point $p$, respectively.

We consider only cases where the slope of $p$ is in the range $0 \le m \le 1$. All other instances can be transformed to this case using rotation and reflection. We begin by fitting the smallest box centered on a point $p$ that intersects both $q$ and $r$. We break the problem into two cases:
\begin{tightenumerate}
 \item The \textit{opposite} case (see figure \ref{fig:conflict-resolution-even}) is where $w^y > 0$, so each box intersects $q$ and $r$ at its top-left and bottom-right corners, respectively.
 \item In the \textit{adjacent} case (see figure \ref{fig:conflict-resolution-adjacent-even}), $w^y < 0$, so the line intersections are adjacent at the top-left and bottom-left corners of the box.
\end{tightenumerate}

@@

Because conflict cell resolution only considers two facets at a time, we may have to iterate multiple times if more than two facets intersect a given cell. If new sample points were found then we add them to the current set $S$ of sample points and return to building the octree from points (section \ref{sec:build-initial-octree}). We finish when the only conflicts identified are at the maximum depth.

@@

Let $M = |F|$ and $N = |V|$, where $F$ are the object facets and $V$ are the object vertices. Let $D$ be the depth of the octree. In this analysis we assume sufficient parallel units to maximize parallelization.

\vspace{2mm}
\paragraph{Time complexity}

>>>>

\documentclass[final,3p,times,twocolumn]{elsarticle}
%\documentclass[submission]{gmp2017}
%\documentclass{gmp2017}

% Put your submission number here.
% This number will appear instead of authors and affiliations if the
% class option "submission" is active.
%\SubNumber{19}

% Put the author names here.
% Use the second argument as a reference to the list of affiliations.
% No authors and affiliations will appear if the class option "submission"
% is active.
%\author{Nathan Vollmer}{1}
%\author{John Edwards}{1}

% Put the affiliations of the authors here.
%\affiliation{1}{Department of Informatics and Computer Science, Idaho State University, USA}

@@

%\usepackage[inline]{showlabels}

@@

%\title{Parallel Quadtree Construction on Collections of Objects}
\title{Fast Construction of Inter-Object Space Representations}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author{}

\address{}

@@

\begin{abstract}
%We present a parallel quadtree algorithm that resolves between geometric objects. Our quadtree has the property that no cell intersects more than one labeled object. A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori, because the number of cells required to resolve closely spaced objects exceeds available memory. Previous parallel algorithms that are spatially adaptive, or that discretize finely in only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation. Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time). We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D. Our results show significant improvement over the current state of the art.
We present a parallel quadtree algorithm that resolves between geometric objects, modeling space between objects rather than the objects themselves. Our quadtree has the property that no cell intersects more than one labeled object. A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori or because the number of cells required to resolve closely spaced objects exceeds available memory. Previous parallel algorithms that are spatially adaptive, discretizing finely only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation. Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time). We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D. Our results show significant improvement over the current state of the art.

@@

%-------------------------------------------------------------------------------
% introduction
%-------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
Constructing quadtrees on objects is an important task with applications in collision detection, distance fields, robot navigation, shape modeling, object description, and other applications. Quadtrees built on objects most often model the objects themselves, providing a space-efficient representation of arbitrarily complex objects. However, our work centers on using quadtrees to separate, or resolve, collections of closely spaced objects, i.e., to construct a discretization such that no cell intersects more than one object. Such quadtrees can be thought of as modeling the space between objects.
% Our quadtree representation is not purposed for fast retrieval as is often the case in hierarchical subdivisions of point data.
% Rather, the quadtrees are used for constructing distance fields, detecting collisions and computing the GVD \cite{edwards2015approximating}.

Modeling inter-object spacing is computationally straightforward when the spacing is large compared to the world bounding box. Approaches typically involve a uniform grid of the space, which leads to efficient computation that often uses graphics processors.

Difficulties arise when objects are close together relative to the size of the domain. An approach using a uniform grid would have excessive memory requirements in order to resolve between objects because the uniformly sized grid cell must be small enough to fit between objects at every location in the domain. Thus, an adaptive approach must be used for datasets of closely spaced objects.

@@

This work extends the work done by Edwards \etal \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm that is adaptive and independent of object distribution. Our algorithm, which is targeted for the GPU, performs an order of magnitude faster than the previous work and will be an important base for later distance transform and generalized Voronoi diagram computation.

@@

Modeling object separation is of some use in 2D (e.g. path planning), but it is a very important problem in many 3D applications. Hierarchically subdividing space between faceted objects in a principled parallel way is complex, and this paper lays the groundwork for our continuing efforts in 3D.

@@

\paragraph{Parallel} Many recent works on fast quadtree construction using the GPU are limited to point sites \cite{bedorf2012sparse,karras2012maximizing,zhou2011data}.
%Kim and Liu's work~\shortcite{kim2014exact} is similar, computing the quadtree on the barycenters of triangles, giving an approximation of an object-resolving quadtree.
 Most quadtree approaches that support surfaces \cite{baert2013out,crassin2009gigavoxels,laine2011efficient,lefebvre2007compressed} are designed for efficient rendering, and actual construction of the quadtree is implemented on the CPU. Two works \cite{bastos2008gpu,park2010cuda} implement Adaptive Distance Fields in parallel on quadtrees but building the quadtree itself is done sequentially.  Yin \etal~\shortcite{yin2011fast} compute the octree entirely on the GPU using a bottom-up approach by initially subdividing into a complete quadtree, resulting in memory usage that is no better than using a uniform grid.
The most similar work to what we do here is Kim and Liu's method~\shortcite{kim2014exact}, which computes the quadtree on the barycenters of triangles, giving an approximation of our quadtree, but without fully resolving between objects.
We have found no GPU quadtree construction method that is fully adaptive and can resolve between objects.

@@

%Our first step is to build a quadtree on the vertices. This step gives us our first approximation to our final quadtree. We use Karras' algorithm \cite{karras2012maximizing} which sorts the Morton codes of the vertices in parallel, then constructs the binary radix tree in parallel. With the binary radix tree, the quadtree can be constructed with a single parallel call. The strength of this algorithm lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points. That is, even if a large number of vertices are clustered in a small area, requiring deep quadtree subdivision, only a constant number of parallel calls need be made.

\subsection{Pruning the quadtree}
\label{sec:pruning}
During initial quadtree construction, we can prune the quadtree to simplify conflict detection and reduce our memory footprint. Assume we have a numeric vertex labeling such that each vertex is labeled to match the object it belongs to. The binary radix tree (BRT) provided by Karras serves as a bounding volume hierarchy and is used to generate the initial quadtree by seperating vertices regardless of their label. Since our objective is to resolve between objects of different labels, we can proactively prune the initial BRT, and subsequently the initial quadtree (see figure \ref{fig:pruning}) such that a leaf node can contain multiple vertices as long as they all have the same label.

%The initial quadtree resolves between vertices, regardless of whether they are labeled differently from each other or not, so each leaf node is guaranteed to contain no more than one vertex. Since our objective is to resolve between objects of different labels, we can proactively prune the initial vertex quadtree (see figure \ref{fig:pruning}) such that a leaf node can contain multiple vertices as long as they all have the same label. By pruning the initial quadtree we simplify conflict detection and reduce our memory footprint.

%\red{Clarify this passage: We remove internal binary radix tree nodes seperating identically labeled vertices. For each binary radix tree leaf in parallel, we first look at the label of the leaf's corresponding vertices. If the colors match, we apply that color to the leaf. Otherwise, we mark that leaf as "required".} This initial coloring can be done immediately after the Karras binary radix tree construction without the need to invoke an additional kernel. Before generating the binary radix tree's corresponding quadtree, we propagate the binary radix leaf colors up the tree in parallel, marking internal nodes as "required" when the colors of two children nodes mismatch. Finally, we generate quadtree nodes from only the required internal binary radix tree nodes. By pruning the initial quadtree, we simplify conflict detection and reduce our memory footprint.

\red{Talk to Nate}
To prune the initial BRT efficiently, we label each BRT node $C$ using the following criterion: if $C$ is a leaf node that separates two vertices with identical labels, label $C$ to match the label of the vertices being separated. If $C$ is a leaf node that separates two vertices having mismatched colors, label $C$ as ``required''. Lastly, if $C$ is an internal node, i.e., it has children, mark it as ``unknown''. This initial step can be done immediately after the Karras BRT construction without the need to invoke an additional kernel.

We then propagate the BRT labels up the tree in parallel, marking ``unknown'' nodes as ``required'' when the labels of the current node's two child nodes don't match. Finally, we generate quadtree nodes from only the required internal binary radix tree nodes.

@@

We define a bounding cell (BCell) to be the smallest internal quadtree node which entirely contains a given facet. Given a facet defined by $n$ endpoints $P=\{p_1, p_2, \dots, p_n\}$, the quadtree address of the BCell is the longest common prefix of the Morton codes of the points in $P$. If a given LCP is more specific than any quadtree node, i.e., if the LCP lies within a quadtree leaf, we simply take the quadtree address of the leaf that the LCP lies within. This is often the case with pruned quadtrees where entire facets may lie within a quadtree leaf. \red{Please check to make sure I got the changes right.} Figure \ref{fig:data-structures-1} gives the addresses of the BCells of the facets in figure \ref{fig:conflict-find-4}.

@@

Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (figure \ref{fig:data-structures-2}).% \texttt{BCells} array construction and sorting is done in parallel with the initial Karras quadtree construction.

@@

%    \node[anchor=south west,inner sep=0,outer sep=0] at (0,-4) {
%%      \subfloat[][]{
%%        \label{fig:conflict-find-2}
%%        \includegraphics[width=0.44\columnwidth]{conflict-find-2.pdf} } \hfill
%%      \subfloat[][]{
%%        \label{fig:conflict-find-4}
%%        \includegraphics[width=0.44\columnwidth]{conflict-find-4.pdf} } \hfill
%      \subfloat[][]{
%        \label{fig:conflict-find-5}
%        \includegraphics[width=0.44\columnwidth]{conflict-find-5.pdf} }
%      \subfloat[][]{
%        \label{fig:conflict-find-6}
%        \includegraphics[width=0.44\columnwidth]{conflict-find-6.pdf} }
%    };

@@

    % two column, .4 size
    % small rectangle
    \draw[red,thick] (0.0,1.65) rectangle (2.33,4.0);
%    % upper-right line
%    \draw[red,thin] (3.27,2.25) -- (3.9,2.25);
%    % lower-left line
%    \draw[red,thin] (2.0,1.05) -- (3.9,0.4);
%    % large rectangle
%    \draw[red,thick] (3.9,0.4) rectangle (5.8,2.25);
  \end{tikzpicture}

%      \subfloat[][]{
%        \label{fig:conflict-find-2}
%        \includegraphics[width=0.44\columnwidth]{conflict-find-2.pdf} } \hfill
%      \subfloat[][]{
%        \label{fig:conflict-find-4}
%        \includegraphics[width=0.44\columnwidth]{conflict-find-4.pdf} } \hfill
      \subfloat[][]{
        \label{fig:conflict-find-5}
        \includegraphics[width=0.44\columnwidth]{conflict-find-5.pdf} }
      \subfloat[][]{
        \label{fig:conflict-find-6}
        \includegraphics[width=0.44\columnwidth]{conflict-find-6.pdf} }

%      \subfloat[][]{
%        \label{fig:data-structures-1}
%        \includegraphics[width=0.38\columnwidth]{data-structures-1.pdf} } \hfill
%      \subfloat[][]{
%        \label{fig:data-structures-2}
%        \includegraphics[width=0.38\columnwidth]{data-structures-3.pdf} } \hfill
%      \subfloat[][]{
%        \label{fig:data-structures-3}
%        \includegraphics[width=0.38\columnwidth]{data-structures-4.pdf} }
  \caption{
%    \protect\subref{fig:conflict-find-1} We have three objects, blue, red, and green with facets labeled A-I.
    \red{Is the pruning correct?}
    We have three objects, blue, red, and green with facets labeled A-I.
    \protect\subref{fig:conflict-find-2} Initial pruned vertex quadtree.
    \protect\subref{fig:conflict-find-4} Zoomed-in to the region outlined by red in \protect\subref{fig:conflict-find-2} and showing the boundary cell (BCell) computation for each facet. %These pairs are given in figure \protect\subref{fig:data-structures-1}.
    \protect\subref{fig:conflict-find-5} Conflict cells, which intersect more than one object, are highlighted.
    \protect\subref{fig:conflict-find-6} The new quadtree after conflict resolution.
%    \protect\subref{fig:data-structures-1} The bounding cells (BCells) are stored in an array initially sorted on facet index (letters are used here for clarity). The quadtree array elements are structures which store child and parent pointers (``C/P'' in the figure).
%    \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access. We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given quadtree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
%\protect\subref{fig:data-structures-3} For a given quadtree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
  }
  \label{fig:conflict-find}
\end{figure}

\begin{figure}
  \centering
      \subfloat[][]{
        \label{fig:data-structures-1}
        \includegraphics[width=0.32\columnwidth]{data-structures-1.pdf} }
      \subfloat[][]{
        \label{fig:data-structures-2}
        \includegraphics[width=0.32\columnwidth]{data-structures-3.pdf} }
      \subfloat[][]{
        \label{fig:data-structures-3}
        \includegraphics[width=0.32\columnwidth]{data-structures-4.pdf} }
  \caption{
    \protect\subref{fig:data-structures-1} The bounding cells (BCells) are stored in an array initially sorted on facet index (letters are used here for clarity). The quadtree array elements are structures which store child and parent pointers (``C/P'' in the figure).
    \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access. We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given quadtree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
\protect\subref{fig:data-structures-3} For a given quadtree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
  }
  \label{fig:data-structures}
\end{figure}

Then we use the \texttt{BCells} array and quadtree data structure to find the conflict cells using algorithm \ref{alg:find-conflict-cells}. We process each leaf cell $L$ in parallel (line \ref{alg:L}). First, we set $L$'s color to $-1$ (uninitialized). We then investigate each ancestor $A$ of $L$ (line \ref{alg:A}) by using the \texttt{Parent} field in the quadtree data structure. Using the \texttt{FFacet} and \texttt{LFacet} fields, we find, respectively, the first and (inclusive) last of possibly multiple facets bounded by $A$ (line \ref{alg:i}). The \texttt{FacetMap} array is used to find all facets bounded by bounding cell $A$ (line \ref{alg:f}). Any facet $f$ for which $A$ is the bounding cell could potentially intersect the leaf cell $L$. We test for intersection between $f$ and $L$ and store the first two facets of differing color (lines \ref{alg:7}-\ref{alg:16}). If at the conclusion of execution $L$.color is equal to $-2$ then $L$ is a conflict cell and must be resolved.

@@

\begin{figure}
  \centering
  \subfloat[][]{
    \label{fig:conflict-resolution-x}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-x.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-even}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-even.pdf} }\\
  \subfloat[][]{
    \label{fig:conflict-resolution-all}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-all.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-quadtree}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-octree.pdf} }\\
  \subfloat[][]{
    \label{fig:conflict-resolution-adjacent-even}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-adjacent-even.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-adjacent-all}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-adjacent-all.pdf} }
  \caption{
    \protect\subref{fig:conflict-resolution-x} A conflict cell with two lines from different objects.
    \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least one sample (red dots).
    \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least two samples. This ensures that a quadtree built from the samples using Karras' algorithm (panel \protect\subref{fig:conflict-resolution-quadtree}) will have no leaf cells that intersect both lines, ensuring that the new quadtree is locally free of conflict cells.
  }
  \label{fig:conflict-resolution}
\end{figure}

@@

\begin{align}
q(t) &= q = q_0 + tv \label{eqn:q} \\
r(f) &= r = r_0 + fw \label{eqn:r}
\end{align}
along with a line
\begin{align}
p(s) &= p = p_0 + su \label{eqn:p}
\end{align}
that bisects $q$ and $r$. Our strategy will be to sample points $P$ on $p(s)$ (figure \ref{fig:conflict-resolution-quadtree}) such that a quadtree built on $S \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent leaf of $C$ will intersect both $q$ and $r$. We do this by ensuring that $P$ is sampled such that every box that intersects both $q$ and $r$ also intersects at least two points in $P$. Because Karras' algorithm guarantees that every leaf cell intersects at most one point, we know that no leaf cell will intersect $q$ and $r$ and thus no leaf cell will be a conflict cell. We will find a series of boxes such that each box's left-most intersection with $p(s)$ is a sample point meeting the above criterion. In the following discussion, $p^x$ and $p^y$ refer to the $x$ and $y$ coordinates of point $p$, respectively.

We consider only cases where the slope of $p$ is in the range $0 \le m \le 1$. All other instances can be transformed to this case using rotation and reflection. We begin by fitting the smallest box centered on a point $p$ that intersects both $q$ and $r$. We break the problem into two cases:
%\begin{tightenumerate}
\begin{enumerate}
 \item The \textit{opposite} case (see figure \ref{fig:conflict-resolution-even}) is where $w^y > 0$, so each box intersects $q$ and $r$ at its top-left and bottom-right corners, respectively.
 \item In the \textit{adjacent} case (see figure \ref{fig:conflict-resolution-adjacent-even}), $w^y < 0$, so the line intersections are adjacent at the top-left and bottom-left corners of the box.
%\end{tightenumerate}
\end{enumerate}

@@

Because conflict cell resolution only considers two facets at a time, we may have to iterate multiple times if more than two facets intersect a given cell. If new sample points were found then we add them to the current set $S$ of sample points and return to building the quadtree from points (section \ref{sec:build-initial-quadtree}). We finish when the only conflicts identified are at the maximum depth.

\subsection{Optimality}
\red{Check this}
Define an optimal quadtree to be one in which only conflict nodes have children, and let an optimal quadtree's size be $n$ total nodes. Our iterative sampling algorithm results in a quadtree that has a size within a constant factor of $n$ in the worst case (see figure \ref{fig:optimality}). We omit the proof as well as an average case analysis because optimality can be achieved by simply performing one final pruning step (see section \ref{sec:pruning}).

\begin{figure}
  \centering
  \subfloat[][]{
    \label{fig:optimality-good}
    \includegraphics[width=0.25\columnwidth]{optimality-good.pdf} }
  \subfloat[][]{
    \label{fig:optimality-bad}
    \includegraphics[width=0.25\columnwidth]{optimality-bad.pdf} }
  \subfloat[][]{
    \label{fig:optimality-bad-all}
    \includegraphics[width=0.25\columnwidth]{optimality-bad-all.pdf} }
  \caption{
    Best and worst cases given two lines. The same number of conflict resolution samples are generated regardless of where the lines are located.
    \protect\subref{fig:optimality-good} Base case: two lines can be resolved by a single quadtree subdivision.
    \protect\subref{fig:optimality-bad} Worst case: the same two lines translated slightly in $y$ now require five subdivisions to be resolved.
    \protect\subref{fig:optimality-bad-all} The number of cells generated from the shown resolution samples is within a constant factor of the worst case.
  }
  \label{fig:optimality}
\end{figure}

@@

Let $M = |F|$ and $N = |V|$, where $F$ are the object facets and $V$ are the object vertices. Let $D$ be the depth of the quadtree. In this analysis we assume sufficient parallel units to maximize parallelization.

%\vspace{2mm}
%\paragraph{Time complexity}
\subsubsection{Time complexity}
